---
title: "Catalysing Mathematical, Computing, and Statistical Innovation Towards Applied Sciences"
author: "Yefeng Yang, Manman Liu, Malgorzata Lagisz, and Shinichi Nakagawa"
date: "`r Sys.Date()`"  # Automatically inserts the current date
output: 
  html_document:
    toc: true             # Adds a table of contents
    toc_float: true       # Makes the TOC float as you scroll
    toc_depth: 3          # Shows up to third-level headers (###)
    number_sections: true # Numbers the sections
    code_folding: show    # Hides code by default, allows toggling
    theme: cosmo          # Bootstrap theme for a modern look
    highlight: tango      # Syntax highlighting style
    df_print: paged       # Pagination for large data frames
    fig_caption: true     # Adds captions to figures
    fig_width: 7          # Default figure width
    fig_height: 5         # Default figure height
    keep_md: true         # Keeps the markdown file after knitting
    smooth_scroll: true   # Smooth scrolling between sections
---

# Required packages

Load packages that are necessary to the data wrangling and modelling. 

```{r, warning=FALSE}
suppressMessages({
  library(dplyr)
  library(readr)
  library(tidyr) 
  library(tidyverse)
  library(stringr)
  library(ggplot2)
  library(ggridges)
  library(ggdist)
  library(metafor)
  library(here)
  library(ggExtra)
  library(boot)
  library(DescTools)
  library(ggsci)
  library(paletteer)
  })
# function
source(here("Func","func.R"))
```


# Load data

Load data bibliometric data from 161 top-ranked mathematics journals (by their 2023 Journal Impact Factor).

```{r, warning=FALSE}
# load
data <- read.csv(here("dat","data.csv"))
# remove na
data <- na.omit(data)
# recode
data <- data %>%
  mutate(Subfield = case_when(
    Subfield == "MATHEMATICS" ~ "Pure Math",
    Subfield == "MATHEMATICS, APPLIED" ~ "Applied Math",
    Subfield == "COMPUTER SCIENCE, ARTIFICIAL INTELLIGENCE" ~ "Artificial Intelligence",
    Subfield == "STATISTICS and PROBABILITY" ~ "Statistical Science",
    Subfield == "COMPUTER SCIENCE, THEORY and METHODS" ~ "Computational Math",
    Subfield == "MATHEMATICS, INTERDISCIPLINARY APPLICATIONS" ~ "Interdisciplinary Math",
    Subfield == "OPERATIONS RESEARCH and MANAGEMENT SCIENCE" ~ "Management Math",
    Subfield == "SOCIAL SCIENCES, MATHEMATICAL METHODS" ~ "Social Math",
    Subfield == "MATHEMATICAL and COMPUTATIONAL BIOLOGY" ~ "Biological Math",
    Subfield == "PHYSICS, MATHEMATICAL" ~ "Physics Math",
    Subfield == "LOGIC" ~ "Logic Math",
    Subfield == "PSYCHOLOGY, MATHEMATICAL" ~ "Psychological Math",
    TRUE ~ Subfield  
  ))
# relevel
levels = c("Pure Math",
           "Applied Math",
           "Physics Math", 
           "Statistical Science",
           "Interdisciplinary Math",
           "Logic Math", 
           "Computational Math",
           "Artificial Intelligence",
           "Biological Math", 
           "Management Math",
           "Social Math",
           "Psychological Math")
data$Subfield <- factor(data$Subfield, levels = levels)
```

# Computation

Calculate the first metric to quantify the percent of citation to top-ranked mathematics journals originating from applied disciplines.

```{r, warning=FALSE}
# subset of "citing journals are non-mathematical journals"
dat <- data %>% filter(Mathematics_group == "No") %>%
  group_by(Journal_cited, Subfield) %>%
    summarize(
    x = sum(Citations, na.rm = TRUE),
    Publisher = first(Publisher),
    Impact_factor = first(Impact_factor),
    Subfield = first(Subfield),
    n = first(N),
    .groups = "drop"
  )

# find the journal with 0 citation from applied fields
out <- setdiff(data$Journal_cited, dat$Journal_cited)
dat.out <- data %>% filter(Journal_cited %in% out) %>%
  group_by(Journal_cited) %>%
    summarize(
    x = 0, # 0 citation from applied fields
    Publisher = first(Publisher),
    Impact_factor = first(Impact_factor),
    Subfield = first(Subfield),
    n = first(N),
    .groups = "drop"
  )

# bind
dat <- rbind(dat, dat.out)

# manual calculation of effect sizes
dat$pi <- with(dat, x/n)
# the logit transformed proportions - log odds
dat <- escalc(measure = "PLO", xi = x, ni = n, data = dat)
```

# Modelling

Let's use the random-effects model and mixed-effects logistic model to get the estimate.

```{r, warning=FALSE}
# normal-normal model
res <- rma(yi, vi, method = "REML", mods = ~ I(Subfield) - 1, test = "knha", data = dat) # remove -1 to get R2 = 54.11%
summary(res)
inv.logit(res$beta)

# sensitivity analysis using logistic regression model with a random intercept
res2 <- rma.glmm(measure = "PLO", xi = x, ni = n, mods = ~ I(Subfield) - 1, data = dat)
res2
inv.logit(res2$beta)
```

# Plot

## Accessibility index

```{r, warning=FALSE}
# update `Subfield` as factor with a reversed levels
levels2 = rev(levels)
dat$Subfield <- factor(dat$Subfield, levels = levels2)
# prepare `dat_fig` with the same level ordering
dat_fig <- dat[, c("Subfield", "pi")]
dat_fig$Subfield <- factor(dat_fig$Subfield, levels = levels2)


# model estimates
dat_fig2 <- data.frame(Subfield = levels, # need to use levels rather than levels2
                        est = inv.logit(res$beta) %>% as.numeric(),
                        lower = inv.logit(res$ci.lb) %>% as.numeric(),
                        upper = inv.logit(res$ci.ub) %>% as.numeric())


# plot
p <- dat_fig %>%
  ggplot(aes(x = Subfield, y = pi, fill = Subfield))  +
  stat_slabinterval(
    aes(slab_alpha = after_stat(f)), 
    thickness = 1, justification = 0.5, color = NA, interval_alpha = 1,
    .width = c(0.66, 0.95)
  ) +
  geom_linerange(
    data = dat_fig2, 
    aes(x = Subfield, ymin = lower, ymax = upper),
    size = 2,
    color = "black"
  ) +
  geom_point(
    data = dat_fig2, 
    aes(x = Subfield, y = est),
    size = 4, shape = 21, color = "black", fill = "white"
  ) +
  geom_text(
    data = dat_fig2,
    aes(x = Subfield, y = est, label = paste0(round(est * 100, 0), "%")), 
    vjust = -0.8,
    size = 3, 
    color = "gray20"
  ) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2), labels =  scales::percent_format(accuracy=2)) + 
  scale_fill_manual(values = paletteer_d("dutchmasters::milkmaid", 12, type = "discrete")) + 
  theme_bw() +
  guides(fill = "none", color = "none", size = "none") +
  theme(legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank()
        ) +
  labs(x = "Subfield of Mathematics", y = "Translational impact") +
  coord_flip()

p
```

## Interdisciplinary connectivity 

```{r, warning=FALSE}
# data frame
dat_fig <- select(data, Subfield, ESI_category)
# filter out ESI mathematics
dat_fig <- dat_fig %>% filter(ESI_category != "Mathematics")
# filter out rows with empty ESI_category
dat_fig <- dat_fig %>%
  filter(ESI_category != "")

# add the first variable in alluvial plot
dat_fig$Mathematics_cited <- rep("Mathematics", nrow(dat_fig))
# plot
datlong_fig <- dlong(dat_fig, Mathematics_cited, Subfield, ESI_category)
datlong_fig$node <- as.factor(datlong_fig$node)
datlong_fig$node <- factor(
  datlong_fig$node,
  levels = c(
    "Psychological Math",
    "Social Math",
    "Management Math",
    "Biological Math",
    "Artificial Intelligence",
    "Computational Math",
    "Logic Math",
    "Interdisciplinary Math",
    "Statistical Science",
    "Physics Math",
    "Applied Math",
    "Pure Math",
    "Social Sciences",
    "Psychiatry and Psychology",
    "Plant and Animal Science",
    "Physics",
    "Pharmacology and Toxicology",
    "Neuroscience and Behavior",
    "Multidisciplinary",
    "Molecular Biology and Genetics",
    "Microbiology",
    "Mathematics",
    "Materials Sciences",
    "Immunology",
    "Geosciences",
    "Environment and Ecology",
    "Engineering",
    "Economics and Business",
    "Computer Science",
    "Clinical Medicine",
    "Chemistry",
    "Biology and Biochemistry",
    "Agricultural Sciences"
  )
)

colors <- c(
  "white",    # White
  "#1f78b4",  # Blue
  "#33a02c",  # Green
  "#e31a1c",  # Red
  "#ff7f00",  # Orange
  "#6a3d9a",  # Purple
  "#b15928",  # Brown
  "#a6cee3",  # Light Blue
  "#b2df8a",  # Light Green
  "#fb9a99",  # Light Red
  "#fdbf6f",  # Light Orange
  "#cab2d6",  # Light Purple
  "#ffff99",  # Light Yellow
  "#8dd3c7",  # Aqua
  "#ffffb3",  # Pale Yellow
  "#bebada",  # Lavender
  "#fb8072",  # Coral
  "#80b1d3",  # Sky Blue
  "#fdb462",  # Peach
  "#b3de69",  # Lime Green
  "#fccde5",  # Pink
  "#bc80bd",  # Light Purple-Pink
  "#ccebc5",  # Light Mint Green
  "#ffed6f",  # Soft Yellow
  "#8c564b",  # Earthy Brown
  "#e377c2",  # Pastel Pink
  "#7f7f7f",  # Neutral Gray
  "#17becf",  # Cyan
  "#aec7e8",  # Light Sky Blue
  "#9edae5",  # Pale Aqua
  "#9467bd",  # Medium Purple
  "#d62728",  # Strong Red
  "#2ca02c"   # Bright Green
)
p <- ggplot(datlong_fig, aes(x = x, next_x = next_x, node = node, next_node = next_node, fill = factor(node), label = node)) +
  sankey_p(flow.alpha = 0.6, node.color = "transparent") + 
  #sankey_p_label(size = 3, color = "white", fill = "gray10", alpha = 0.6) + 
  sankey_p_label(size = 4, color = "gray10", fill = colors, alpha = 0.9) +
  #scale_fill_manual(values = wes_palette("Cavalcanti1", n = 5)) +
  #scale_fill_aaas() +
  theme_sankey(base_size = 10) +
  labs(x = NULL, title = " ") +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 14, hjust = .5),
        axis.text.x = element_text(color = "black", size = 14)) +
  scale_x_discrete(labels = c(" ", "Subfield of Mathematics", "Applied field of science"), position = "top")

p
```



